---
title: "eda_template"
author: "Aaron Grzasko"
date: "May 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE, comment=NA,
                      options(width=120), message=FALSE,
                      warning=FALSE, fig.align='center')
```

```{r}
# load libraries and custom functions

if (!require(pacman)) install.packages('pacman'); library(pacman)

p_load(ggplot2, dplyr, ggthemes, geoR,stringr, knitr, readr, moments, DataExplorer)

source('https://raw.githubusercontent.com/datafeelings/CUNY_DATA_621_SS2018_Final/master/2_code/Graphical_Functions.R')

source('https://raw.githubusercontent.com/datafeelings/CUNY_DATA_621_SS2018_Final/master/2_code/stats_function.R')

```  
  
```{r}

# Load the full data
inp <- read_csv('https://raw.githubusercontent.com/datafeelings/CUNY_DATA_621_SS2018_Final/master/0_raw_data/bcn_listings.csv.gz')
inp <- data.frame(inp)

# fix price variable
inp$price <- as.numeric(gsub("[\\$,]", "", inp$price))

```
  
## Variable Overview  

### host_since  
  
```{r}
# convert to date; then conver to number of years since minimum date
inp$host_since <- as.Date(inp$host_since)
inp$host_since <- as.numeric(inp$host_since - min(inp$host_since, na.rm=T)) / 365.25
```  
  
#### Distribution  
```{r}
stats(data.frame(inp$host_since))
```  
  
#### Histogram  
```{r}  
hist621(inp, host_since)
```  
  
#### Box-Cox  
  
```{r}
boxcoxfit(inp$host_since[!is.na(inp$host_since)] + 0.001)

```
  
#### Relationship to Price  
  
```{r}
scat621(inp, host_since^1.2, log(price))

```  
  

  
### host_about  
  
```{r}
# convert to length
inp$host_about <- nchar(inp$host_about)
inp$host_about[is.na(inp$host_about)] <- 0

```  
  
#### Distribution  
```{r}
stats(data.frame(inp$host_about))

```

#### Histogram  
  
```{r}
hist621(inp, host_about)
```  
  
#### Box-Cox  
  
```{r}
boxcoxfit(inp$host_about + 1)
```
  
#### Relationship to Price  
  
```{r}
scat621(inp, log(host_about+1), log(price))

```  
  
### host_response_time  
  
#### Distribution  
```{r}
inp$host_response_time <- factor(inp$host_response_time)

table(inp$host_response_time, useNA = "always")

group_category(data=inp,feature="host_response_time", threshold=0.0001)

```
  
#### Bar Chart  
  
```{r}

bar621(inp,
       reorder(host_response_time, -table(host_response_time)[host_response_time]),
       xlab="host_response_time",
       title="host_response_time by type") + ylim(c(0,10050))
```
  
#### Relationship to Price  
  
```{r}
box621(inp, host_response_time, log(price))

```
  
### host_response_rate  
  
```{r}
inp$host_response_rate <- as.numeric(str_replace(inp$host_response_rate, "%" ,""))/100
```  
  
#### Distribution  
  
```{r}
stats(data.frame(inp$host_response_rate))


```
  
#### Histogram  
  
````{r}
hist621(inp, host_response_rate, bins=15)

```
  
```{r}
boxcoxfit(inp$host_response_rate[!is.na(inp$host_response_rate)]+0.001)

```
  
#### Relationship to Price  
  
```{r}
scat621(inp, host_response_rate^4, log(price))

```  
  
### host_is_superhost  
  
#### Distribution  
````{r}
table(inp$host_is_superhost, useNA = 'always')
````  
  
#### Bar Chart  
  
```{r}
bar621(inp,host_is_superhost) +  ylim(c(0,17000))

```  
  
#### Relationship to Price 
    
```{r}
box621(inp,host_is_superhost,log(price))  
  
```  
  
### host_listings_count  
  
#### Distribution  
  
```{r}
stats(data.frame(inp$host_listings_count))
```  
  
#### Histogram  
  
```{r}
hist621(inp, host_listings_count)

```  
  
#### Box Cox  
  
```{r}
boxcoxfit(inp$host_listings_count[!is.na(inp$host_listings_count)]+0.001)

```  
  
#### Relationship to Price  
  
```{r}
scat621(inp,log(host_listings_count),log(price))

```  
  
### host_identity_verified  
  
#### Distribution  
  
```{r}
inp$host_identity_verified <- factor(inp$host_identity_verified)
table(inp$host_identity_verified, useNA = "always")
```  
  
#### Bar Chart  
  
```{r}
bar621(inp,host_identity_verified) + ylim(c(0,10500))

```  
  
#### Relationship to Price  
  
```{r}
box621(inp, host_identity_verified, log(price))

```  
  
### property_type  
  
```{r}
inp$property_type <- factor(inp$property_type)

```
  
#### Distribution  
  
```{r}
sort(table(inp$property_type, useNA = "always"), decreasing = T)

```  
  
#### Bar Chart  
  
Original  
```{r}
bar621(inp, reorder(property_type, -table(property_type)[property_type]),
       vadj = -50, hadj = 0.2,title="property_type by type",xlab="property_type") + ylim(c(0,18000)) + theme(axis.text.x = element_text(angle = 90, hjust = 1))


```  
  
#### Relationship to Price  
  
```{r}
box621(inp,reorder(property_type, -table(property_type)[property_type]),
       log(price),
       title = "log(price) vs. property_type", xlab="property_type") + theme(axis.text.x = element_text(angle = 90, hjust = 1))


```
