---
title: "eda_template"
author: ""
date: "5/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exploratory Analysis

```{r}
# Load the required packages
library(ggplot2)
# library(caret)
library(e1071)
library(dplyr)
library(plyr)
library(readr)
library(boot)
library(car)
# library(DataExplorer)
library(knitr)
library(MASS)


# Load the stats and vis functions
source("2_code/Graphical_Functions.R")
source("2_code/stats_function.R")

```

```{r}
# Load the full data

inp = read_csv("0_raw_data/bcn_listings.csv.gz")
inp = data.frame(inp)
```

```{r}
# Variable subset and preprocess

newdata = subset(inp, select = c(price,review_scores_communication,review_scores_location,review_scores_value,requires_license,instant_bookable,is_business_travel_ready,cancellation_policy,require_guest_profile_picture,require_guest_phone_verification,calculated_host_listings_count,reviews_per_month,neighbourhood_cleansed,zipcode,neighbourhood_group_cleansed))

# Changing factors to int char
newdata$price = as.numeric(gsub("[\\$,]", "", newdata$price))


```

### Variable overview


**review_scores_communication**
  
*Distribution*
  
```{r}
kable(stats(data.frame(newdata$review_scores_communication)))
```

*Histogram*

```{r}
tmp = "review_scores_communication"
bar621(df=newdata, review_scores_communication, xlab="", title=paste("Histogram of",tmp))
```



**review_scores_location**
  
*Distribution*
  
```{r}
kable(stats(data.frame(newdata$review_scores_location)))
```

*Histogram*

```{r}
tmp = "review_scores_location"
bar621(df=newdata, review_scores_location, xlab="", title=paste("Histogram of",tmp))
```

**review_scores_value**
  
*Distribution*
  
```{r}
kable(stats(data.frame(newdata$review_scores_value)))
```

*Histogram*

```{r}
tmp = "review_scores_value"
bar621(df=newdata, review_scores_value, xlab="", title=paste("Histogram of",tmp))
```


**requires_license**

Variable excluded as all observations have the same value.
  
  
**instant_bookable**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(table(newdata$instant_bookable,dnn = "instant_bookable"))
```

*Relative frequency*

```{r}
tmp = "instant_bookable"
bar621(df=newdata, instant_bookable, xlab="", title=paste("Frequency of levels of",tmp))
```


**is_business_travel_ready**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(table(newdata$is_business_travel_ready,dnn = "is_business_travel_ready"))
```

*Relative frequency*

```{r}
tmp = "is_business_travel_ready"
bar621(df=newdata, is_business_travel_ready, xlab="", title=paste("Frequency of levels of",tmp))
```


**cancellation_policy**
    
This is a factor with five levels corresponding to cancellation policy types.
  
*Distribution*
  
```{r}
kable(table(newdata$cancellation_policy,dnn = "cancellation_policy"))
```

*Relative frequency*

```{r}
tmp = "cancellation_policy"
bar621(df=newdata, cancellation_policy, xlab="", title=paste("Frequency of levels of",tmp))
```


**require_guest_profile_picture**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(table(newdata$require_guest_profile_picture,dnn = "require_guest_profile_picture"))
```

*Relative frequency*

```{r}
tmp = "require_guest_profile_picture"
bar621(df=newdata, require_guest_profile_picture, xlab="", title=paste("Frequency of levels of",tmp))
```


**require_guest_phone_verification**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(table(newdata$require_guest_phone_verification,dnn = "require_guest_phone_verification"))
```

*Relative frequency*

```{r}
tmp = "require_guest_phone_verification"
bar621(df=newdata, require_guest_phone_verification, xlab="", title=paste("Frequency of levels of",tmp))
```

**calculated_host_listings_count**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(stats(data.frame(newdata$calculated_host_listings_count)))
```

*Histogram*

```{r}
tmp = "calculated_host_listings_count"
hist621(df=newdata, calculated_host_listings_count, xlab="", title=paste("Histogram of",tmp))
```

**reviews_per_month**
    
This is a factor with two levels: true ("t") and false ("f")
  
*Distribution*
  
```{r}
kable(stats(data.frame(newdata$reviews_per_month)))
```

*Histogram*

```{r}
tmp = "reviews_per_month"
hist621(df=newdata, reviews_per_month, xlab="", title=paste("Histogram of",tmp))
```

**neighbourhood_cleansed**
    
This is a factor with two 72 levels correspondent to the neighborhoods of Barcelona. Will not be considered further as data on districts the neighborhoods belong to is available (see below).
  
*Distribution*
  
```{r}
str(data.frame(newdata$neighbourhood_cleansed))
```
  

**neighbourhood_group_cleansed**
    
This is a factor with ten levels correspondent to the ten districts of the city (location of the listing).
  
*Distribution*
  
```{r}
kable(table(newdata$neighbourhood_group_cleansed,dnn = "neighbourhood_group_cleansed"))
```

*Relative frequency*

```{r}
tmp = "neighbourhood_group_cleansed"
bar621(df=newdata, neighbourhood_group_cleansed, xlab="", title=paste("Frequency of levels of",tmp))
```
