---
title: "final_project_initial"
author: "Dmitriy Vecheruk"
date: "4/23/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(caret)
library(DataExplorer)
library(AppliedPredictiveModeling)
library(vcd)
library(gridExtra)
```


```{r message=F,warning=F}


inp = read_csv("0_raw_data/bcn_listings.csv") 
```

```{r vis_functions include=F}
plot_overview = function(df, x) {
df = data.frame(df)
y = df[,x]

p1 = ggplot(df, aes(x= x, y = y)) + geom_boxplot() + xlab(x) + ylab("value") + 
  ggtitle(x)
p2 = ggplot() + geom_density(aes(x=y),na.rm = T) + 
  geom_vline(aes(xintercept=mean(y,na.rm = T)), color = "red",linetype = 2) + xlab(x) +
   ggtitle("")
p3 = ggplot() + geom_point(aes(sample=y), stat = "qq", size = 0.1)
p4 = ggplot()  + geom_density(aes(x=log(y)),na.rm = T) + xlab(paste0("log(",x,")"))

grid.arrange(p1, p2, p3, p4, nrow = 2)
}

plot_target_cor = function(df,x,trg_var_name){

df_cc = df[,c(x,trg_var_name)]
df_cc = as.data.frame(df_cc[complete.cases(df_cc),]) 
trg_cor = cor.test(df_cc[,x],df_cc[,trg_var_name])
df_cc$x_log = log(df_cc[trg_var_name>0,x])
trg_log_cor = cor.test(df_cc[df_cc$x_log>0,"x_log"],df_cc[df_cc$x_log>0,trg_var_name])

cor_out = paste("cor:",round(trg_cor$estimate,3),
                "p-val:", ifelse(trg_cor$p.value<0.01,"<0.01",round(trg_cor$p.value,3)))
cor_out_log = paste("cor (log):",round(trg_log_cor$estimate,3),
                "p-val:", ifelse(trg_log_cor$p.value<0.01,"<0.01",round(trg_log_cor$p.value,3)))

x_pos = quantile(df_cc[,x],0.9)
y_pos_0 = quantile(df_cc[,trg_var_name],0.006)
y_pos_1 = quantile(df_cc[,trg_var_name],0.004)

p = ggplot(data = df_cc, aes_string(x=x,y=trg_var_name)) +
  geom_point(size=1, alpha = 0.5) + geom_smooth(method = "lm") +
  annotate("text", x = x_pos, y = y_pos_0,label = cor_out,size=3) +
  annotate("text", x = x_pos, y = y_pos_1,label = cor_out_log, size=3) +
  ylab(trg_var_name) + ggtitle(paste(x,"vs",trg_var_name))
p
}

plot_overview_target = function(df,x,trg_var_name){
  grobs = list(plot_overview(df,x),plot_target_cor(df,x,trg_var_name))
  grid.arrange(grobs = grobs, nrow = 1,widths=c(5,4), 
               top = paste(x,"overview and correlation vs.",trg_var_name))
}
```

### Inspect the variable subset

```{r}
var_subset = c("price", "review_scores_communication","review_scores_location","review_scores_value","requires_license","instant_bookable","is_business_travel_ready","cancellation_policy","require_guest_profile_picture","require_guest_phone_verification","calculated_host_listings_count","reviews_per_month","neighbourhood_cleansed","zipcode")

inp_subset = inp %>% select(one_of(var_subset))

```

#### Summary 

```{r cache=T}

dollars_to_numeric = function(input) {
  out = sub("\\$", "", input)
  out = as.numeric(sub(",", "", out))
  return(out)
}

inp_subset= inp_subset %>% 
  mutate_at(c("price"),dollars_to_numeric) %>% 
  mutate_at(c("requires_license","instant_bookable","is_business_travel_ready","cancellation_policy",
              "require_guest_profile_picture","require_guest_phone_verification",
              "neighbourhood_cleansed","zipcode","review_scores_communication",
              "review_scores_location","review_scores_value"),as.factor)
summary(inp_subset)
```

#### Missing values

```{r}
plot_missing(inp_subset, title = "Percentage of missing data per variable")
```

1) Variables with missing data: `zipcode`(5%), Review values `reviews_per_month`(20%),`review_scores_communication`(21%),
`review_scores_location`(21%),`review_scores_value`(21%) - these missing observations are for the same rows

#### NZV

```{r }
nzv = nearZeroVar(inp_subset, saveMetrics= TRUE)
nzv
```

1) zero variance - kick out: `requires_license`: 
2) near-zero variance - unlikely to be useful: `is_business_travel_ready`, `require_guest_profile_picture`,`require_guest_phone_verification`

#### Distributions of discrete variables

```{r}
# Drop the index column
df = inp_subset %>% select(-requires_license)

df_viz = split_columns(data = df)
```

```{r}
## View distribution of all discrete variables
plot_bar(df_viz$discrete, title = "Frequency counts for discrete variables")
```

1) Too many levels for a non-binned factor: `zipcode`: 99 categories, `neighbourhood_cleansed`: 72 categories - can probably cut down digits of zip and impute missing based on neighborhood average zip!
2) The three review scores  variables are very unbalanced and should be transformed to 10,9,8, and below 8.
3) `cancellation_policy`: merge both "super_strict" groups into one level (or even into "strict")

#### Distributions of continuous variables

```{r}
## View distribution of all continuous variables
plot_histogram(df_viz$continuous, "Histograms of continous variables")
```

1) Price is log-distributed

```{r}
plot_histogram(log(df$price),"Histogram of log(price)")
```

2) `calculated_host_listings_count` is Poisson-distributed (**Q:** How to transform it?)

```{r}
plot_histogram(df$calculated_host_listings_count,"Histogram of calculated_host_listings_count")
```

3) `reviews_per_month` is very skewed and can be log-transformed

```{r}
plot_histogram(log(df$reviews_per_month),"Histogram of log(reviews_per_month)")
```

#### 1.2. Bivariate analysis
  
```{r}
## View correlation of all continuous varaibles (for complete cases)

plot_correlation(df_viz$continuous[complete.cases(df_viz$continuous),],
                 type = "continuous",title = "Pairwise correlations between continuous variables")
```


The pairwise relationships between the continuous variables and the response are displayed below

```{r}
qplot(calculated_host_listings_count,price,data=df,alpha=0.5) + ggtitle("calculated_host_listings_count vs price")
```

```{r}
qplot(calculated_host_listings_count,log(price),data=df,alpha=0.5) +
  geom_smooth(method = "lm")+
  ggtitle("calculated_host_listings_count vs log(price)")

```

1) A positive link with the price for `calculated_host_listings_count`, also can try binning!

```{r}
qplot(reviews_per_month,price,data=df, alpha=0.5) +
  ggtitle("reviews_per_month vs price")
```

```{r}
qplot(log(reviews_per_month),log(price),data=df,alpha=0.5) + 
  geom_smooth()+
  ggtitle("Log-log: reviews_per_month vs price")

# qplot(reviews_per_month,log(price),data=df %>% filter(reviews_per_month<50),alpha=0.5) + ggtitle("Log-log: reviews_per_month vs price")
```

2) No relationship with `reviews_per_month`.

####  Discrete predictors

```{r}
df1 = cbind(df_viz$discrete, log(df$price))

disc_vars = names(df_viz$discrete)

for (i in disc_vars){
  plot_boxplot(data = df1,by = i,notch=T)
}

```

1) Useful predictors: `is_business_ready`(t:+), `cancellation_policy`(strict:+),
`require_phone_verification`(t:+), probably `neighborhood` or `zip`
