---
title: "Models"
author: "Georgia Galanopoulos"
date: "May 25, 2018"
output: html_document
---

```{r}
library(Amelia)
library(MASS)
library(pROC)
library(Metrics)
library(car)
library(gvlma)

train_inp = read.csv("https://raw.githubusercontent.com/datafeelings/CUNY_DATA_621_SS2018_Final/master/2_code/train_inp.csv", header = T)
test_inp = read.csv("https://raw.githubusercontent.com/datafeelings/CUNY_DATA_621_SS2018_Final/master/2_code/test_inp.csv", header = T)
```

Selecting only the variables that are significant (classified as 2).
```{r}
two = train_inp[ , -which(names(train_inp) %in% c("host_response_rate","host_is_superhost","host_identity_verified",
                                                "room_type","accommodates","beds","extra_people","review_scores_communication",
                                                "review_scores_location","review_scores_value","instant_bookable",
                                                "require_guest_profile_picture", "reviews_per_month", "maximum_nights"))]
```

To determine handling NA values in review_scores variables, we take two approaches: remove NA values or remove the variables themselves.


# Remove NA rows
```{r}
#####################################
############ MODEL 1 ################
#####################################
#### Remove NA rows
subr = subset(two, !is.na(review_scores_rating)&!is.na(review_scores_accuracy)&!is.na(review_scores_cleanliness)&!is.na(review_scores_checkin))
norows = stepAIC(lm(price~ ., data=subr), direction = "both")
summary(norows)

test_inp$values = predict(norows, test_inp)
#plot(multiclass.roc(test_inp$price, test_inp$values), print.auc = TRUE)

qqPlot(norows, main="Residuals QQ Plot")
hist(studres(norows), main="Distribution of Studentized Residuals")
vif(norows)
spreadLevelPlot(norows)
summary(gvlma(norows))
```


# Remove Columns
```{r}
#####################################
############ MODEL 2 ################
#####################################
#### Remove review columns
subc = two[ , -which(names(two) %in% c("review_scores_rating", "review_scores_accuracy", "review_scores_cleanliness", "review_scores_checkin"))]
nocolumns = stepAIC(lm(price~., data=subc), direction = "both")
summary(nocolumns)
rmse_noc = sqrt(mean((nocolumns$residuals)^2))

test_inp$values2 = predict(nocolumns, test_inp)
#plot(multiclass.roc(test_inp$price, test_inp$values2), print.auc = TRUE)


qqPlot(nocolumns, main="Residuals QQ Plot")
hist(studres(nocolumns), main="Distribution of Studentized Residuals")
vif(nocolumns)
spreadLevelPlot(nocolumns)
summary(gvlma(nocolumns))
```


Removing the rows was the model that performed better. But that also means that we will have to remove any NA values from the test set if we are to use it.
```{r}
missmap(subset(test_inp, select = c(price, values, values2)))
```
